---
title: "github_document"
author: "Christopher Crowe"
date: "September 13, 2018"
output: html_document
---



```{r library chunk, include = FALSE}
library(tidyverse)
```

# Problem 1

Here is a code chunk that creates a dataframe with the following elements:
 
  * a random sample of size 10 from a uniform[0,5] distribution
  * a logical vector indicating whether elements of the sample are greater than 2
  * a (length-10) character vector
  * a (length-10) factor vector
  
```{r sample chunk}
hw_df = tibble(
  random_sample_10 = runif(10, min = 0, max = 5),
  vec_logical_gt2 = random_sample_10 > 2,
  vec_char_10 = c("This","is", "a", "character", "vector", "with", "a", "length", "of", "ten."),
  vec_factor_10 = c("male","male","female", "female", "female", "male", "male", "female", "female", "male")
)
```

The following code chunk attempts to calculate the mean of each of the four elements in the dataframe, hw_df:

```{r first mean attempt chunk}
mean(hw_df$random_sample_10)
mean(hw_df$vec_logical_gt2)
mean(hw_df$vec_char_10)
mean(hw_df$vec_factor_10)
```

The output of the above code chunk shows that we are able to calculate a mean  of approximately `r round(mean(hw_df$random_sample_10), digits = 2)` and `r round(mean(hw_df$vec_logical_gt2), digits = 2)` for the **random sample** and the **logical vector**, respectively. 

However, we are _not_ able to calculate the mean of the **character vector** and the **factor vector**. The associated error message explains that this is because the mean() function requires numeric or logical arguments. Since the character vector and factor vector are neither numeric nor logical, the mean() function returns `r mean(hw_df$vec_char_10)` for these two vectors.

The following code chunk shows how we can attempt to calculate the means of the logical, character, and factor vectors by treating them as numeric types. 

```{r second mean attempt chunk, eval = FALSE}
mean(as.numeric(hw_df$vec_logical_gt2))
mean(as.numeric(hw_df$vec_char_10))
mean(as.numeric(hw_df$vec_factor_10))
```

The output of the above code chunk shows that the mean of the logical vector is the same as the mean we calculated during our first attempt (i.e. `r round(mean(hw_df$vec_logical_gt2), digits = 2)`).

However, we are still _unsuccessful_ in our attempt to calculate the means of the character and factor vectors.
