---
title: "github_document"
author: "Christopher Crowe"
date: "September 13, 2018"
output: github_document
---



```{r library chunk, include = FALSE}

## load the library tidyverse
library(tidyverse)
```

# Problem 1

Here is a code chunk that creates a dataframe with the following elements:
 
  * a random sample of size 10 from a uniform[0,5] distribution
  * a logical vector indicating whether elements of the sample are greater than 2
  * a (length-10) character vector
  * a (length-10) factor vector
  
```{r sample chunk}

## set up the data frame
hw_df = tibble(
  random_sample_10 = runif(10, min = 0, max = 5),
  vec_logical_gt2 = random_sample_10 > 2,
  vec_char_10 = c("This","is", "a", "character", "vector", "with", "a", "length", "of", "ten."),
  vec_factor_10 = factor(c("male","male","female", "female", "female", "male", "male", "female", "female", "male"))
)

```

The following code chunk attempts to calculate the mean of each of the four elements in the dataframe, hw_df:

```{r first mean attempt chunk}

## take the mean of the random sample
mean(hw_df$random_sample_10)

## take the mean of the logical vector
mean(hw_df$vec_logical_gt2)

## take the mean of the character vector
mean(hw_df$vec_char_10)

## take the mean of the factor vector
mean(hw_df$vec_factor_10)
```

The output of the above code chunk shows that we are able to calculate a mean  of approximately `r round(mean(hw_df$random_sample_10), digits = 2)` and `r round(mean(hw_df$vec_logical_gt2), digits = 2)` for the **random sample** and the **logical vector**, respectively. 

However, we are _not_ able to calculate the mean of the **character vector** and the **factor vector**. The associated error message explains that this is because the mean() function requires numeric or logical arguments. Since the character vector and factor vector are neither numeric nor logical, the mean() function returns `r mean(hw_df$vec_char_10)` for these two vectors.

The following code chunk shows how we can attempt to calculate the means of the logical, character, and factor vectors by treating them as numeric types. 

```{r second mean attempt chunk, eval = FALSE}

## convert logical to numeric and take the mean
mean(as.numeric(hw_df$vec_logical_gt2))

## convert character to numeric and take the mean
mean(as.numeric(hw_df$vec_char_10))

## convert factor to numeric and take the mean
mean(as.numeric(hw_df$vec_factor_10))
```

The output of the above code chunk shows that the mean of the logical vector is the same as the mean we calculated during our first attempt (i.e. `r round(mean(as.numeric(hw_df$vec_logical_gt2)), digits = 2)`). Further, we see that the mean of the factor vector can now be computed and is equal to `r round(mean(as.numeric(hw_df$vec_factor_10)), digits = 2)`.

However, we are still _unsuccessful_ in our attempt to calculate the mean of the character vector because treating the character vector as numeric results in NA.

The following code chunk shows what happens when we try to calculate the mean after performing the following conversions: character to factor to numeric and factor to character to numeric.

```{r third mean attempt chunk}

## convert character to factor to numeric and take the mean
mean(as.numeric(as.factor(hw_df$vec_char_10)))

## convert factor to character to numeric and take the mean
mean(as.numeric(as.character(hw_df$vec_factor_10)))
```

The above output shows that we are able to calculate the mean of the character vector after converting it from character to factor to numeric (i.e. `r round(mean(as.numeric(as.factor(hw_df$vec_char_10))), digits = 2)`). However, we are now unable to calculate the mean for the factor vector that has been converted from factor to character to numeric. 

# Problem 2

The following code chunk creates a dataframe with the following elements:

  * x: a random sample of size 1000 from a standard Normal distribution
  * y: a random sample of size 1000 from a standard Normal distribution
  * a logical vector indicating whether x + y > 0
  * a numeric vector created by coercing the above logical vector
  * a factor vector created by coercing the above logical vector
  
```{r data frame chunk}

## set up data frame
hw_df_2 = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  xy_positive = (x + y) > 2,
  logical_as_numeric = as.numeric(xy_positive),
  logical_as_factor = as.factor(xy_positive)
)
```
  
The dataframe created by the above code chunk has a size of: rows = `r nrow(hw_df_2)`, columns = `r ncol(hw_df_2)`. The mean of x is `r round(mean(hw_df_2$x), digits = 2)`. The median of x is `r round(median(hw_df_2$x), digits = 2)`. The proportion of cases for which the logical vector is TRUE is `r mean(hw_df_2$xy_positive)`.

The following code chunk creates a 3 different scatter plots: 
  * one colored using the logical vector
  * one colored using the logical vector that has been treated as numeric
  * one colored using the logical vector that has been treated as a factor 

```{r scatterplot chunk}

## create plot colored using logical vector and save plot in project directory
ggplot(hw_df_2, aes(x = x, y = y, color = hw_df_2$xy_positive)) + geom_point()
ggsave("myplot.pdf")

## create plot colored using logical vector as numeric
ggplot(hw_df_2, aes(x = x, y = y, color = hw_df_2$logical_as_numeric)) + geom_point()

## create plot colored using logical vector as factor
ggplot(hw_df_2, aes(x = x, y = y, color = hw_df_2$logical_as_factor)) + geom_point()
```

The above scatter plots show how the color scales differ depending on the properties of the variables used in the color statement. Specifically, since the logical and factor vectors have discrete values, there are discrete color codes. On the other hand, using a numeric variables results in a gradient of colors because a numeric variable can be continuous. 

